How to Stop Useless Refactorings

### 2016-08-13 Manuscript.tex
Updates from ShareLaTeX

Note that these managers are {\em not} asking for better
automatic refactoring
tool support. Their human programmers are refactoring
code all the time, perhaps too much so.
Hence, their management goal is: 
\bi
\item
{\em Not} to optimize {\em necessary} refactorings;
\item
But rather to stop {\em unnecessary} refactorings.
\ei
For those managers, we offer the  following heuristic:
\bi
\item If there is no evidence in the historical record that changing ``X'' is useful;
\item Then stop developers from doing ``X''.
\ei
This paper describes one way to operationalize this heuristic, in the context
of assessing bad smells. We use data mining to generate two oracles: a {\em primary
change oracle} and a {\em secondary validation oracle}.
By combining these two oracles,
 we can generate and validate useful
changes to code modules.

@@

As to 
the {\em secondary validation oracle}, this  checks if the primary oracle is proposing
sensible changes. For this purpose we  use Random Forests\cite{XXX}, augmented with SMOTE (synthetic  minority over-sampling technique~\cite{YYY}).
 In our framework,  learning
the secondary oracle is   a {\em separate} task to learning the primary
oracle. This  ensures that the validation oracle offers an independent
opinion on the value of the proposed changes.

When these oracles were applied to defect data from
open source object-oriented software systems, we discovered a {\em conjunctive fallacy}
in the bad smell literature~\cite{erni96,bender99,Shatnawi10,Alves2010,hermans15}.
A common heuristic is to recommend refactoring of ``bad'' code
to make it ``better''   as follows: for all static code measures that exceed some threshold, make changes such that the thresholds are no longer exceeded. That is:
\begin{equation}\label{eq:df}
\scriptstyle
\begin{array}{rl}
\mathit{bad} & = \left(a_1 > t_1 \right) \bigvee \left(a_2 > t_2\right) \bigvee    ... \\
\mathit{better} & = \neg\;\mathit{bad} = \left(a_1 \le t_1 \right) \bigwedge \left(a_2 \le t_2\right)  \bigwedge  ...
\end{array}
\end{equation}
We say that the above definition of ``better'' is a conjunctive fallacy
since it assumes that the best way to improve code is to decreases multiple code attribute measures below $t_i$ in order
to remove the ``bad'' smells. This is wrong since 
there exists an intricate set of associations between static code measures
such that {\em decreasing}  $a_i$ necessitates {\em increasing} $a_j$.
It is easy to see why this is so.
A requirement for  refactoring is that the new code supports the same functionality
as before. If we pull code out of a function (since that function has grown too
large), {\em that functionality has to go somewhere else}. What we see in XTREE's change
recommendation is that if we decrease a module's lines of code  then in the refactored
code, the functionality is achieved via a communication between several different parts
of the system. That is, {\em decreasing} the lines of code metric within a module {\em increases}
that module's coupling metrics.

  The rest of this paper is structured as follows. First, we offer some background on bad smells. After that,
  in Section~3, 
  different methods for determining when code does/does not have bad smell. One of those methods
  will be XTREE while the others will come from other researchers. Next,    Section~4
  defines an assessment method that checks which of these methods is most useful. The results from that assessment 
  are presented in Section~5. Those results will answer five research questions.

 {\bf  RQ1: Effectiveness}: According to the validation oracle, which of the methods  defined in Section~3 is the best  change oracle for identifying what and how
  code modules should be changed? To answer this question, we used data from five OO Java projects
  (Ivy, Lucene, Ant, Poi, Jedit). It was found that:
  \begin{lesson}
 XTREE is the best oracle on how to change code modules (in order to reduce defects).
  \end{lesson}

@@

{\bf RQ5: Conjunctive Fallacy}:  
Is it always  useful  to aplly \eq{df}; i.e. make code better by   reducing the values of  multiple code attributes? We find that:
\begin{lesson}
XTREE usually recommends reducing lines of code (size of the modules).
That said,  XTREE often recommends {\em increasing} the values of other static code attributes.
\end{lesson} 
Note that {\bf RQ3, RQ4, RQ5} 
confirms the intuitions
of the project managers that prompted this investigation:
\bi
\item
Yes,  indeed, their programmers
were wasting time on   performing refactorings  that  decreased multiple measures when, in fact,
they should be trying to {\em decrease} some measures while {\em increasing} others.
\item
Across all studied projects, XTREE found that  around one-fifth  
  of the measures were usually found within  change recommendations. That is, it is 80\% likely that
  that refactoring based on reducing any one measurement picked at random will be useful.
\ei
Consequently, we  recommend that XTREE be used as follows:
\bi
\item Before doing  refactoring based on a bad smell report...
\item ...check and stop any refactoring   for which there is no proof
  in the historical record that that change improves the code.
\ei
Note that this use case applies to both manual refactorings proposed by developers
as well as refactoring as conducted by automatic methods~\cite{mkaouer2015many}.
Note also that
this recommendation assumes that there is an historical record. If none exists, then the results of this paper could be use used as a guide (see our Table 8).

 \subsection{Replication Package}

  In order to support dissemination
  of research results, and to encourage other researchers to repeat our results,
  all the code used in this paper is publicly available\footnote{XXX: need a public repo}.

  \subsection{Relationship to Prior Work }
A four page prelininary report on the XTREE system~\cite{krishna2015actionable} has been presented previously\footnote{https://goo.gl/2In3Lr}. That short report offered case studies only two  of
the five data sets studied here. Also, that prior work had:
\bi
\item
no comparison to any baseline
method from other researchers, whereas this paper compares XTREE's recommendations
with those found by researchers exploring bad smells.
\item
no {\em secondary validation oracle}, i.e. no way to check if XTREE's recommended changes were sensible.
\ei
Further, of this entire paper, the only sections containing material
found in prior papers is 3.2.1, 3.2.2 as well as two-fifths of \fig{}.
%YYY section1,2, 3.1,4,
% Prior to XTREE, the authors worked on
% the CD algorithm discussed in ~\tion{cdcd}. 
% The goal of CD was the same as WHERE-- find recommendations on how to best change software
% modules. However, as discussed in~\tion{cdcd},
% CD worked very different to WHERE. Further, as shown in this paper,
% CD is significantly out-performed by XTREE.

@@

Having made the case for automatic support for assessing bad smells,
this section reviews different ways for building those tools (one of those tools,
XTREE, will be our recommended {\em primary change oracle}).
Later in this paper, we will offer  a {\em secondary validation oracle}
that checks the effectiveness of the changes proposed by XTREE.

@@

\subsubsection{Enri \& Lewerentz}
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentzz~\cite{erni96} found the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.
Shatnawi and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$ since it does not consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.

@@

\input{tex/prediction_stats.tex}      
\section{Evaluation}
The previous section proposed numerous methods for detecting bad smells that need
to be resolved. This section offers an experiment of the following form:
\bi
\item Using each method as a {\em primary change oracle} to recommend how code should be changed.
\item Apply those changes. 
\item Run a {\em secondary validation oracle} to assess the defect-proness of the changed code.
\item Sort the change oracles on how well they reduce defects (as judged by the validation oracle).
\ei
Using this experiment, we can address the research questions dicsussed in the introduction.

{\bf RQ1: Effectiveness: } which of the methods defined in Section 3 is the best change oracle for identifying what and how code modules should be changed? 
To   answer to this question, we will assume that developers
refactor their code until the bad smell thresholds is not violated.
This refactoring will start with some {\em initial} code
base that is changed to a {\em new} code base. 
For example, if the bad smell is \mbox{{\em loc $>100$}} and a 
code module has 500 lines of code, we reason
optimistically that we can change that code metric
to 100.  
Using the secondary validation oracle,  we then predict the
number of defects in $d_+,d_-$ in {\em initial} and {\em new}. We evaluate the performance of XTREE, CD, Shatnawi, and Alves methods in setting bad smell thresholds.
The best bad smell threshold method is the one that maximizes
\begin{equation}\label{eq:diff}
\mathit{improvement} = 100* \left(1 - \frac{ d_- }{ d_+}\right)
\end{equation}

@@

\subsection{Secondary Validation Oracles}
\label{sect:eval}

As mentioned in the introduction, our proposed framework has two oracles:
a primary change oracle (XTREE) and the secondary validation oracle described in this section.

@@

To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a {\em secondary validation oracle} (which is learned separately
to the primary oracle) that   assesses
how defective is the code before and after some
refactoring. 
For their second oracle,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}.
A shortcoming of QMOOD
is that quality models learned from other projects
may perform poorly when applied to new projects~\cite{localvsglobal}.
Hence, for this study, we  eschew
older quality models like QMOOD. Instead, we use
Random Forests~\cite{Breiman2001} to learn defect predictors
from OO code metrics.
Note that, unlilke QMOOD, such measurements 
  are specific to the project, and the measurements can be reconstructed and the predictors rebuilt for other projects.

@@

\fig{counts} counted how {\em often} XTREE's recommendations mentioned a static code attribute.
\fig{changes}, on the other hand, shows the {\em direction} of the recommended change:
\bi
\item Gray bars show  an  {\em increase} to a static code measure;
\item White bars shows a   {\em drecrease} to a static code measure;
\item Bars that are all white or all gray indicate that in our 40 repeated runs, XTREE recommended changing an attribute the same way, all the time.
\item Bars that are mixtures of white and gray mean that, sometimes, XTREE makes different recommendations about how to change a static
code attribute.
\ei
Based on \fig{changes}, we say that {\bf Result4} is: the direction of change recommended  XTREE is  very stable repeated runs of the program  (evidence:
the bars are mostly the same color).

@@

{\em  Is  it  always  useful  to  apply \eq{df};  i.e.   make  code  better  by  reducing  the  values  of multiple code attributes?

Note that, in \fig{changes}, the bars are colored white and gray; i.e. XTREE recommends {\em decreasing} and {\em increasing} 
static attribute values. That is, always decreasing static code measures (as suggested by \eq{df}) is {\em not} recommended by XTREE.

To explore this point further, we conducted the following ``what-if'' study. Once XTREE builds its trees with its leaf clusters then:
\be
\item For all clusters $C_0$, $C_1$ where the percent of defects in $C_0$ is greater than $C_1$...
\item For all attribute measures that are have a statistically significantly different distribution  between $C_0$ and $C_1$...
\item Report the direction of the change.
\ee
Note that the changes found in this way are somewhat more general than those found by XTREE since XTREE's recommendations are limited
to comments on the test set given to the program. The above three points comments on the direction of {\em all possible changes}
that XTREE could ever report

Figure ~\ref{fig:multcomp} shows the results of this ``what if'' analysis. As in  \fig{changes}, there many increases
and decreases. Also, only with  cbo, rfc, loc,cam is a change recommended for a majority of the
data sets  (three of five, or more). 
XXXX

ups and down. fucntionalty has to go somehwere else

highlights all possible changes proposed by XTREE, upon close inspection it becomes apparent that XTREE suggests that it is perhaps better to reduce only certain forms of coupling. Refactoring operations that reduce one form of coupling at the cost of increasing another can sometimes be beneficial. This is most noticeable in Ivy, where one of the better approaches is to reduce coupling between objects (cbo) (efferent coupling) while increasing the afferent coupling (CA), again implying a redistribution of responsibilities. Using thresholds may not always appreciate the existence of these co-changes between metrics, and as a result many refactoring operations are usually left unexplored.

Note that we prefer XTREE to Shatnawi and the Alves
method since those other methods
do not reason
2about decreases {\em and} increases of multiple 
code metrics, at the same time.

As a final comment, we note
that  \fig{changes} enables us to explain the uniformity
of the results seen with Jedit in \fig{jur}.
Observe how in \fig{changes} the only change ever
found is a reduction to {\em rfc}. Clearly, in this
data set, there is very little that can be usefully changed.

\fig{multcomp} shows the frequency with which the methods
recommend changes to specific code methods.
Note that XTREE proposes thresholds to
few code metrics compared to the other approaches.

When combining  \fig{jur}, \fig{counts}, and \fig{multcomp}, we   see that
even though XTREE proposes changes to far fewer code metrics, those few
code metrics are usually just as effective (or
more effective) than the multiple
thresholds
proposed by CD, Shatnawi or Alves.  That is, XTREE proposes
{\em fewer} and better thresholds than the other approaches studied here.

\begin{lesson}
In addition to recommending useful changes, XTREE recommends changes to all co-varying software metrics.
\end{lesson}

\section{Threats to Validity}\label{sect:valid}
XXXX assumes anything can change

assumes we can get data

assuems that defect is a;l;l.

Before  that, we offer one following note on a
core assumption of XTREE.
XTREE   is focused on Fowler's bad smells
as contributors to software defects. Some researchers 
 suggest that discussing bad smells is useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about code smells and refactorings can (a)~introduce new comers to the code base or (b)~share knowledge on a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects.

As with any empirical study, several issues can affect the final results. Therefore, any
conclusions made from this work must be offered with the following
caveats.

\subsection{External Validity}
Based on the experimental results above,
as well as the discussion in \tion{prelim},
we believe that bad smell indicators (e.g. \mbox{{\em loc}$>$ 100})
have limited external validity beyond the projects from which they are derived. 
While specific models are externally valid,
there may still be general methods like XTREE for finding the good local models.

Our definition of bad smells is limited to those represented by OO code metrics (a premise often used in related work).   
XTREE, Shatnawi, Alves et al. can  only comment
on bad smells   expressed as code metrics 
  in the historical log of a project.

If developers want to justify their refactorings
via bad smells expressed in other terminology,
then the  analysis of this paper must:
\begin{itemize}
    \item Either wait till 
data about those new
terms has been collected. 
\item Or, apply cutting edge transfer learning
methods~\cite{Nam15,Jing15} to map data from other projects
into the current one.
\end{itemize}
Note that the transfer learning approach would
be highly experimental and require more study
before it can be safely recommended.

Sampling bias threatens any data mining experiment; i.e., what matters
there may not be true here. For example, the data sets used here comes from Jureczko et al. and any biases in their selection procedures
threaten the validity of these results. 
That said,
the best we can do is define our methods and publicize our data and code so that other researchers can
try to repeat our results and, perhaps, point out a previously unknown bias
in our analysis. Hopefully, other researchers will emulate our methods in
order to repeat, refute, or improve our results.

% Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.  
% That said, we have taken care to document in this paper the decisions made by engineering
% judgement that could effect our conclusions.

@@

\subsection{  Internal Validity}\label{sect:coc}

% The main internal validity threat   in this study was the use of
% optimistic reasoning when generating the changed code modules.
We assumed in \tion{rq}
that any code module can be changed in order
to stop triggering a bad smell detector. This is not necessarily
true, as noted in {\em PROBLEM \#2} in this paper's introduction.
To mitigate this issue:
\begin{itemize}
    \item We do not propose using XTREE as a bad
smell generator;
\item Rather, we offer it as  a critique agent for bad smells
proposed by some other oracle (e.g. a human developer).
\end{itemize}
We showed about that XTREE's recommendations are not verbose.
That is, even if XTREE recommends the superset of possibly important bad smells,
that superset is not so large as to  overlap with many
 bad smell proposed at random 
by   developers.

XTREE does not require a working defect predictor to generate
its bad smell thresholds. However, to {\em certify}
that those thresholds are satisfactory, some
second quality predictor must be applied. A data set containing defect counts before and after refactoring activities would enable an accurate assessment of the performance of the threshold-setting and recommender methods, however, such a data set is not available at this time.

>>>>

Note that these managers are {\em not} asking for better
automatic refactoring
tool support. Their human programmers are refactoring
code all the time, perhaps too much so.
Hence, their management goal is: 
\bi
\item
{\em Not} to optimize {\em necessary} refactorings;
\item
But rather to stop {\em unnecessary} refactorings.
\ei
For those managers, we offer the  following heuristic:
\bi
\item If there is no evidence in the historical record that changing ``X'' is useful;
\item Then stop developers from doing ``X''.
\ei
This paper describes one way to operationalize this heuristic, in the context
of assessing bad smells. We use data mining to generate two oracles: a {\em primary
change oracle} and a {\em secondary verification oracle}.
By combining these two oracles,
 we can generate and validate useful
changes to code modules.

@@

As to 
the {\em secondary verification oracle}, this  checks if the primary oracle is proposing
sensible changes. For this purpose we  use Random Forests\cite{XXX}, augmented with SMOTE (synthetic  minority over-sampling technique~\cite{YYY}).
 In our framework,  learning
the secondary oracle is   a {\em separate} task to learning the primary
oracle. This  ensures that the verification oracle offers an independent
opinion on the value of the proposed changes.

When these oracles were applied to defect data from
open source object-oriented software systems, we discovered a {\em conjunctive fallacy}
in the bad smell literature~\cite{erni96,bender99,Shatnawi10,Alves2010,hermans15}.
A common heuristic is to recommend refactoring of ``bad'' code
to make it ``better''   as follows: for all static code measures that exceed some threshold, make changes such that the thresholds are no longer exceeded. That is:
\begin{equation}\label{eq:df}
\scriptstyle
\begin{array}{rl}
\mathit{bad} & = \left(a_1 > t_1 \right) \bigvee \left(a_2 > t_2\right) \bigvee    ... \\
\mathit{better} & = \neg\;\mathit{bad} = \left(a_1 \le t_1 \right) \bigwedge \left(a_2 \le t_2\right)  \bigwedge  ...
\end{array}
\end{equation}
We say that the above definition of ``better'' is a conjunctive fallacy
since it assumes that the best way to improve code is to decreases multiple code attribute measures below $t_i$ in order
to remove the ``bad'' smells. This is wrong.  
There exists an intricate set of associations between static code measures
such that {\em decreasing}  $a_i$ necessitates {\em increasing} $a_j$.
It is easy to see why this is so.
A requirement for  refactoring is that the new code supports the same functionality
as before. If we pull code out of a function (since that function has grown too
large), {\em that functionality has to go somewhere else}. What we see in XTREE's change
recommendation is that if we decrease a module's lines of code  then in the refactored
code, the functionality is achieved via a communication between several different parts
of the system. That is, {\em decreasing} the lines of code metric within a module {\em increases}
that module's coupling metrics.

\subsection{Structure of this Paper}

This paper  claims that (a)~XTREES is a  better way than \eq{fd}  to define useful changes to software and (b)~those changes are useful for recognizing   superfluous refactorings since usually refer to a very 
small subset of the static code measures (which allows us to rule out refactorings based on any of the omitted
static code measures). 
In support of these claims, 
the rest of this paper is structured as follows. First, we offer some background on bad smells. After that,
  in Section~3, 
  different methods for determining when code does/does not have bad smell. One of those methods
  will be XTREE while the others will come from other researchers. Next,    Section~4
  defines an assessment method that checks which of these methods is most useful. The results from that assessment 
  are presented in Section~5. Those results will answer five research questions.

 {\bf  RQ1: Effectiveness}: According to the verification oracle, which of the methods  defined in Section~3 is the best  change oracle for identifying what and how
  code modules should be changed? To answer this question, we used data from five OO Java projects
  (Ivy, Lucene, Ant, Poi, Jedit). It was found that:
  \begin{lesson}
 XTREE is the best oracle on how to change code modules (in order to reduce defects).
  \end{lesson}

@@

{\bf RQ5: Conjunctive Fallacy}:  
Is it always  useful  to aplly \eq{df}; i.e. make code better by   reducing the values of  multiple code attributes? We find that:
\begin{lesson}
XTREE usually recommends reducing lines of code (size of the modules).
That said,  XTREE often recommends {\em increasing} the values of other static code attributes.
\end{lesson} 
Note that {\bf RQ3, RQ4, RQ5} 
confirms the intuitions
of the project managers that prompted this investigation:
\bi
\item
Yes,  indeed, their programmers
were wasting time on   performing refactorings  that  decreased multiple measures when, in fact,
they should be trying to {\em decrease} some measures while {\em increasing} others.
\item
Across all studied projects, XTREE found that  around one-fifth  
  of the measures were usually found within  change recommendations. That is, it is 80\% likely that
  that refactoring based on reducing any one measurement picked at random will be useful.
\ei
Consequently, we  recommend the following use case for  XTREE:
\bi
\item Before doing  refactoring based on a bad smell report...
\item ...check and stop any refactoring   for which there is no proof
  in the historical record that that change improves the code.
\ei
This use case described  applies to both manual refactorings proposed by developers
as well as refactoring as conducted by automatic methods~\cite{mkaouer2015many}. That is, XTREE could optimize automatic
code refactoring by discouraging refactorings for useless goals.

\subsection{ Threats to Validity}

The results of this paper care biased by out choice of refactoring goal (reducing defects) and our choice
of measures collected from software project (OO measures such as
  depth of inheritance, number of child classes, etc).

  That said, it should be possible extend the methods of this paper to  other kinds of other kinds of goals (e.g. maintainability, reliability, security,
  or the knowledge sharing measures favoured by Bosu and Carver~\cite{bosu13}) and other kinds of inputs (e.g. the process measures favored by Rahman, Devanbu et al.\cite{Rahman2013})  as follows:
  \bi
  \item Find a data source for the other measures of interest;
  \item Implement another secondary verification oracle that can assess maintainability, reliability, security, etc;
\item Implement a better primary verification oracle that can do ``better'' that XTREES at finding changes (where ``better'' is defined in terms 
of the opinions of the verification oracle).
\ei
To assist other researchers exploring these points, we offer a full replication package for this study at  https://github.com/ai-se/XTREE\_IST.

The results of this paper are also specific to the data sets explored here.
Such sampling bias threatens any data mining experiment; i.e., what matters
{\em there} may not be true {\em here}. For example, the data sets used here comes from a survey of
open source JAVA projects from Jureczko et al.~\cite{jureczko10}. Any biases in their selection procedures
threaten the validity of these results.

That said,
the best we can do is define our methods and publicize our data and code so that other researchers can
try to repeat our results and, perhaps, point out a previously unknown bias
in our analysis. Hopefully, other researchers will emulate our methods in
order to repeat, refute, or improve our results.

The recommendation to management, shown on page one of this paper,  was to stop developers doing ``X'' if there was
nothistorical evidence that ``X'' was useful. 
This recommendation assumes that such an  historical record can be accessed. If not,  then the results of this paper could be use used as a guide (see our Table 8). Alternatively, some form of transfer learning~\cite{Nam15,Jing15,krishna16} could be used to import data from another project (but more research would be required to test the validity of combing tools like XTREEs with transfer learning).

  \subsection{Relationship to Prior Work }
A four page prelininary report on the XTREE system~\cite{krishna2015actionable} has been presented previously\footnote{https://goo.gl/2In3Lr}. That short report offered case studies only two  of
the five data sets studied here. Also, that prior work had:
\bi
\item
no comparison to any baseline
method from other researchers, whereas this paper compares XTREE's recommendations
with those found by researchers exploring bad smells.
\item
no {\em secondary verification oracle}, i.e. no way to check if XTREE's recommended changes were sensible.
\ei
Further, of this entire paper, the only sections containing material
found in prior papers is 3.2.1, 3.2.2 as well as two-fifths of the results in \fig{xxx}.

@@

Having made the case for automatic support for assessing bad smells,
this section reviews different ways for building those tools (one of those tools,
XTREE, will be our recommended {\em primary change oracle}).
Later in this paper, we will offer  a {\em secondary verification oracle}
that checks the effectiveness of the changes proposed by XTREE.

@@

\subsubsection{Enri \& Lewerentz}
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentzz~\cite{erni96} found the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.
Shatnawi and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$ since it does not consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical verification.

@@

\input{tex/prediction_stats.tex}      
\section{Evaluation}
The previous section proposed numerous methods for detecting bad smells that need
to be resolved. This section offers an experiment of the following form:
\bi
\item Using each method as a {\em primary change oracle} to recommend how code should be changed.
\item Apply those changes. 
\item Run a {\em secondary verification oracle} to assess the defect-proness of the changed code.
\item Sort the change oracles on how well they reduce defects (as judged by the verification oracle).
\ei
Using this experiment, we can address the research questions dicsussed in the introduction.

{\bf RQ1: Effectiveness: } which of the methods defined in Section 3 is the best change oracle for identifying what and how code modules should be changed? 
To   answer to this question, we will assume that developers
refactor their code until the bad smell thresholds is not violated.
This refactoring will start with some {\em initial} code
base that is changed to a {\em new} code base. 
For example, if the bad smell is \mbox{{\em loc $>100$}} and a 
code module has 500 lines of code, we reason
optimistically that we can change that code metric
to 100.  
Using the secondary verification oracle,  we then predict the
number of defects in $d_+,d_-$ in {\em initial} and {\em new}. We evaluate the performance of XTREE, CD, Shatnawi, and Alves methods in setting bad smell thresholds.
The best bad smell threshold method is the one that maximizes
\begin{equation}\label{eq:diff}
\mathit{improvement} = 100* \left(1 - \frac{ d_- }{ d_+}\right)
\end{equation}

@@

\subsection{Secondary verification Oracles}
\label{sect:eval}

As mentioned in the introduction, our proposed framework has two oracles:
a primary change oracle (XTREE) and the secondary verification oracle described in this section.

@@

To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a {\em secondary verification oracle} (which is learned separately
to the primary oracle) that   assesses
how defective is the code before and after some
refactoring. 
For their second oracle,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}.
A shortcoming of QMOOD
is that quality models learned from other projects
may perform poorly when applied to new projects~\cite{localvsglobal}.
Hence, for this study, we  eschew
older quality models like QMOOD. Instead, we use
Random Forests~\cite{Breiman2001} to learn defect predictors
from OO code metrics.
Note that, unlilke QMOOD, such measurements 
  are specific to the project, and the measurements can be reconstructed and the predictors rebuilt for other projects.

@@

\fig{counts} counted how {\em often} XTREE's recommendations mentioned a static code attribute.
\fig{changes}, on the other hand, shows the {\em direction} of the recommended change:
\bi
\item Gray bars show  an  {\em increase} to a static code measure;
\item White bars shows a   {\em drecrease} to a static code measure;
\item Bars that are all white or all gray indicate that in our 40 repeated runs, XTREE recommended changing an attribute the same way, all the time.
\item Bars that are mixtures of white and gray mean that, sometimes, XTREE makes different recommendations about how to change a static
code attribute.
\ei
Based on \fig{changes}, we say that {\bf Result4} is that the direction of change recommended  XTREE is  very stable repeated runs of the program  (evidence:
the bars are mostly the same color).

\fig{changes} also comments on the inter-relationships between static code attributes. Note that while some measures in
\fig{changes} are decreased, many are increased.  
For example, consider the {\em Poi} results
from \fig{counts} that recommends decreasing {\em loc}
but making large increases to {\em cbo} (coupling between
objects). Here, XTREE is advising us to break up
large classes class by into services
in other classes. Note that such a refactoring will, by
definition, increase the coupling between objects. 
Note also that such increases  to reduce
defects would never be proposed by the outlier methods
of Shatnawi or Alves since their core assumption is that bad
smells arise from unusually large code metrics.

@@

{\em  Is  it  always  useful  to  apply \eq{df};  i.e.   make  code  better  by  reducing  the  values  of multiple code attributes?}

In  \fig{changes}, the bars are colored both white and gray; i.e. XTREE recommends {\em decreasing} and {\em increasing} 
static attribute values. That is, always decreasing static code measures (as suggested by \eq{df}) is {\em not} recommended by XTREE.

To explore this point further, we conducted the following ``what-if'' study. Once XTREE builds its trees with its leaf clusters then:
\be
\item For all clusters $C_0$, $C_1$ where the percent of defects in $C_0$ is greater than $C_1$...
\item For all attribute measures that are have a statistically significantly different distribution  between $C_0$ and $C_1$...
\item Report the direction of the change.
\ee
Note that the changes found in this way are somewhat more general than the above results
since they were  limited
to comments on the test set given to the program. The above three points comments on the direction of {\em all possible changes}
that XTREE could ever report

Figure ~\ref{fig:multcomp} shows the results of this ``what if'' analysis. As might be expected, the recommendation is to
always reduce lines of code (loc). But for the other attributes, there are many recommendations to increase those values.
Hence, {\bf Result5} is that while XTREE always  recommends reducing loc,
it also   often recommends increasing the values of other static code attributes.

% of ``better'' deefined as per \eq{fd} is misleading.

% Also, most attributes are not changed in most data sets.
% Only our attributes (cbo, rfc, loc,cam) change in a majority of data sets  (three of five, or more).

% XXXX

% ups and down. fucntionalty has to go somehwere else

% highlights all possible changes proposed by XTREE, upon close inspection it becomes apparent that XTREE suggests that it is perhaps better to reduce only certain forms of coupling. Refactoring operations that reduce one form of coupling at the cost of increasing another can sometimes be beneficial. This is most noticeable in Ivy, where one of the better approaches is to reduce coupling between objects (cbo) (efferent coupling) while increasing the afferent coupling (CA), again implying a redistribution of responsibilities. Using thresholds may not always appreciate the existence of these co-changes between metrics, and as a result many refactoring operations are usually left unexplored.

% Note that we prefer XTREE to Shatnawi and the Alves
% method since those other methods
% do not reason
% 2about decreases {\em and} increases of multiple 
% code metrics, at the same time.

% As a final comment, we note
% that  \fig{changes} enables us to explain the uniformity
% of the results seen with Jedit in \fig{jur}.
% Observe how in \fig{changes} the only change ever
% found is a reduction to {\em rfc}. Clearly, in this
% data set, there is very little that can be usefully changed.

% \fig{multcomp} shows the frequency with which the methods
% recommend changes to specific code methods.
% Note that XTREE proposes thresholds to
% few code metrics compared to the other approaches.

% When combining  \fig{jur}, \fig{counts}, and \fig{multcomp}, we   see that
% even though XTREE proposes changes to far fewer code metrics, those few
% code metrics are usually just as effective (or
% more effective) than the multiple
% thresholds
% proposed by CD, Shatnawi or Alves.  That is, XTREE proposes
% {\em fewer} and better thresholds than the other approaches studied here.

% \begin{lesson}
% In addition to recommending useful changes, XTREE recommends changes to all co-varying software metrics.
% \end{lesson}

% \section{Threats to Validity}\label{sect:valid}
% XXXX assumes anything can change

% assumes we can get data

% assuems that defect is a;l;l.

% Before  that, we offer one following note on a
% core assumption of XTREE.
% XTREE   is focused on Fowler's bad smells
% as contributors to software defects. Some researchers 
%  suggest that discussing bad smells is useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
% debates about code smells and refactorings can (a)~introduce new comers to the code base or (b)~share knowledge on a team
% about interactions within the code. In that view, 
% the concept of refactoring can be useful
% even if it does not address issues of defects.

% As with any empirical study, several issues can affect the final results. Therefore, any
% conclusions made from this work must be offered with the following
% caveats.

% \subsection{External Validity}
% Based on the experimental results above,
% as well as the discussion in \tion{prelim},
% we believe that bad smell indicators (e.g. \mbox{{\em loc}$>$ 100})
% have limited external validity beyond the projects from which they are derived. 
% While specific models are externally valid,
% there may still be general methods like XTREE for finding the good local models.

% Our definition of bad smells is limited to those represented by OO code metrics (a premise often used in related work).   
% XTREE, Shatnawi, Alves et al. can  only comment
% on bad smells   expressed as code metrics 
%   in the historical log of a project.

% If developers want to justify their refactorings
% via bad smells expressed in other terminology,
% then the  analysis of this paper must:
% \begin{itemize}
%     \item Either wait till 
% data about those new
% terms has been collected. 
% \item Or, apply cutting edge transfer learning
% methods~\cite{Nam15,Jing15} to map data from other projects
% into the current one.
% \end{itemize}
% Note that the transfer learning approach would
% be highly experimental and require more study
% before it can be safely recommended.

% Sampling bias threatens any data mining experiment; i.e., what matters
% there may not be true here. For example, the data sets used here comes from Jureczko et al. and any biases in their selection procedures
% threaten the validity of these results. 
% That said,
% the best we can do is define our methods and publicize our data and code so that other researchers can
% try to repeat our results and, perhaps, point out a previously unknown bias
% in our analysis. Hopefully, other researchers will emulate our methods in
% order to repeat, refute, or improve our results.

% % Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.  
% % That said, we have taken care to document in this paper the decisions made by engineering
% % judgement that could effect our conclusions.

@@

% \subsection{  Internal Validity}\label{sect:coc}

% % The main internal validity threat   in this study was the use of
% % optimistic reasoning when generating the changed code modules.
% We assumed in \tion{rq}
% that any code module can be changed in order
% to stop triggering a bad smell detector. This is not necessarily
% true, as noted in {\em PROBLEM \#2} in this paper's introduction.
% To mitigate this issue:
% \begin{itemize}
%     \item We do not propose using XTREE as a bad
% smell generator;
% \item Rather, we offer it as  a critique agent for bad smells
% proposed by some other oracle (e.g. a human developer).
% \end{itemize}
% We showed about that XTREE's recommendations are not verbose.
% That is, even if XTREE recommends the superset of possibly important bad smells,
% that superset is not so large as to  overlap with many
%  bad smell proposed at random 
% by   developers.

% XTREE does not require a working defect predictor to generate
% its bad smell thresholds. However, to {\em certify}
% that those thresholds are satisfactory, some
% second quality predictor must be applied. A data set containing defect counts before and after refactoring activities would enable an accurate assessment of the performance of the threshold-setting and recommender methods, however, such a data set is not available at this time.

%  Historically,
% SE researchers~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} have used the QMOOD hierarchical quality model~\cite{Bansiya02}.
% Since such quality models may be very project specific~\cite{localvsglobal}, we prefer building local 
% defect predictors using Random Forests. We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.  
% Our results show that when a defect log contains too few
% non-defective examples then defector predictors
% perform poorly (see the data sets in the last few lines of \fig{j})
% as do their associated bad smell threshold generators
% (recall the performance of Lucene and Poi in~\fig{jur}). It should be noted that this is not a limitation of XTREE-- rather
% it seems fundamental to the problem of certifying bad smells.
% As evidence of this, recall that {\em all} the bad smell threshold generators of~\fig{jur}
% had issues with the defect data sets with  many defects
% (Lucene and Poi).
